Formalizing Stack Safety as a Security Property

### 2023-02-03 submission5/core.tex
Testing updates

Indeed, the call stack is a ancient~\cite{one06phrack} and
perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
  DBLP:conf/sp/SzekeresPWS13,
  DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
target for low-level attacks, frequently involving control-flow hijacking via corrupting
the return address, and memory corruption more generally.

@@

There are several ways to evaluate whether an enforcement mechanism enforces the above
stack safety properties. Ideally such validation would be done through formal proof over
the semantics of the enforcement-augmented machine.
However, while there are no fundamental barriers to producing such a proof,
it would be considerable work to carry out for a full ISA like RISC-V and
complex enforcement mechanisms like Roessler and DeHon's micro-policies.
We therefore choose to systematically \emph{test} their micro-policies.

We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
and extend it with a runtime monitor implementing a stack safety
micro-policy. We choose the Coq proof assistant as the setting for our implementation
to ensure that our coinductive trace definitions are well-formed, to reason about
them, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.

To use QuickChick, we build random test-case generators that produce
(1) an initial RISC-V machine state, including the program to be executed;
(2)
  an initial policy state, tagging instructions and the stack regions appropriately
  for the micro-policy (see \cref{subsec:policies})
(3)
  a map from instructions to lists of their operations in the security semantics.

To write such generators we build on the work of
Hri\c{t}cu et al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
introduced {\em generation by execution} to produce progams that lead
machines towards interesting behaviors. We extend the technique with a degree
of statefulness, to nudge the generator toward executions that will carry
on without early failstops and thus have a better chance of reaching a
state that presents opportunities for an attack.

We need to extend typical testing schemes further to handle the nested
nature of confidentiality: rather than just generating two
initial machines that are variants of one another and letting them
execute to test for noninterference, we generate a new variant
{\em every time a call is made} and check confidentiality for the
subtrace produced from that variant state until its corresponding
return. As a result, a ``single'' confidentiality test compactly
checks multiple nested calls.

Our primary testing targets are the eager {\em Depth Isolation}
and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.

>>>>

Indeed, the call stack is an ancient~\cite{one06phrack} and
perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
  DBLP:conf/sp/SzekeresPWS13,
  DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
target for low-level attacks, frequently involving control-flow hijacking via corrupting
the return address, and memory corruption more generally.

@@

There are several ways to evaluate whether an enforcement mechanism enforces the above
stack safety properties. Ideally such validation would be done through formal proof over
the semantics of the enforcement-augmented machine.
However, while there are no fundamental barriers to producing such a proof,
it would be considerable work to carry out for a full ISA like RISC-V and
complex enforcement mechanisms like Roessler and DeHon's micro-policies.
We therefore choose to systematically \emph{test} their micro-policies.
\leo{moved the following here, I think it makes more sense here.}
Our primary testing targets are the eager {\em Depth Isolation}
and the {\em Lazy Per-Activation Tagging and Clearing} micro-policies.

We use a Coq specification of the RISC-V architecture~\cite{Bourgeat2021AMF},
and extend it with a runtime monitor implementing a stack safety
micro-policy. We choose the Coq proof assistant as the setting for our implementation
to ensure that our coinductive trace definitions are well-formed, to reason about
them\leo{where do we mention reasoning about them?}, and to leverage the power of the QuickChick property-based testing framework~\cite{Pierce:SF4}.

\subsection{Test Generation}
To use QuickChick, we developed random test-case generators that produce
(1) an initial RISC-V machine state, including the program to be executed;
(2)
  an initial policy state, tagging instructions and the stack regions appropriately
  for the micro-policy (see \cref{subsec:policies})
(3)
  a map from instructions to lists of their operations in the security semantics.

To write such generators we build on the work of Hri\c{t}cu et
al. \cite{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, which
introduced {\em generation by execution}, a technique that produces
programs that lead to longer executions---and hopefully towards more
interesting behaviors as a result. We extend their technique with an additional
degree of statefulness: when complex calling conventions need to be
observed to avoid early failstops, extra care must be taken to ensure
that any such calls are well behaved. For example, immediately
following a call we need to initialize any stack-allocated variables
with an increased probability. However, at the same time, we need to
allow for the possibility of {\em not} performing such an initialization
to allow for potential attack vectors to manifest.

Moreover, we need to further extend typical hyperproperty testing
schemes to handle the nested nature of confidentiality: rather than
just generating two initial machines that are variants of one another
and letting them execute to test for noninterference, we generate a
new variant {\em every time a call is made} and check confidentiality
for the subtrace produced from that variant state until its
corresponding return. As a result, a ``single'' confidentiality test
compactly checks multiple nested calls.
